\ttitle

### 2015-08-05 Chapters/Introduction.tex
Introduction - Body Composition
- Completed obesity and surgical outcomes

Increased serum bilirubin and bile salts directly affect lymphocyte function as well as the response to pro-inflammatory cytokines. Bile salts in higher than normal concentrations also induce apoptosis in hepatocytes. Biliary obstruction is also associated with significant alterations in hepatic macro and micro-circulation. This in turn can affect the phagocyte clearing function of the liver. Obstructive jaundice is also associated with abnormal cytokine homeostasis predisposing the patient to impaired wound healing, sepsis and increased risk of peri-operative mortality.  For instance, concentrations of TNF-\alpha and IL-6 are significantly elevated in jaundiced patients while the concentrations of IL-1 and IL-2 are decreased.

@@

>>>>

Increased serum bilirubin and bile salts directly affect lymphocyte function as well as the response to pro-inflammatory cytokines. Bile salts in higher than normal concentrations also induce apoptosis in hepatocytes. Biliary obstruction is also associated with significant alterations in hepatic macro and micro-circulation. This in turn can affect the phagocyte clearing function of the liver. Obstructive jaundice is also associated with abnormal cytokine homeostasis predisposing the patient to impaired wound healing, sepsis and increased risk of peri-operative mortality.  For instance, concentrations of TNF-$\alpha$ and IL-6 are significantly elevated in jaundiced patients while the concentrations of IL-1 and IL-2 are decreased.

@@

Obesity is an increasing problem in all developed nations with the proportion of obese patients undergoing surgery increasing every year. 40\% of the adult population in the United Kingdom are expected to have a body mass index of over 30 kg/m$^2$ by 2025. Obesity, especially morbid obesity, is associated with a wide range of secondary disease states such as ischaemic heart disease, diabetes mellitus, hypertension and increased incidence of cancers.

Obesity presents unique problems in surgical patients for several reasons. Surgery and anaesthesia in the morbidly obese patients can be technically challenging. The comorbidities associated with morbid obesity require evaluation and management in the perioperative period. Postoperative physiotherapy and mobilisation require particular attention in the morbidly obese patients.

\subsection{Obesity and Surgical Outcomes}

The impact of obesity on the rate of postoperative complications remains controversial. It has been a commonly held belief that obese patients had greater incidence of complications. Some of the adverse outcomes such as wound complications and respiratory dyfunction where conseidered to be the direct result of the increased body mass index while other complications where related to the associated comorbidities such as cardiovascular disease, obstructive sleep apnea or diabetes mellitus.

Several studies have reported increased incidence of postoperative complications after cardiac, transplant, urologic and gynaecological surgery. \parencite{prem_operative_1965, holley_obesity_1990, fasol_influence_1992, gruberg_impact_2002, lee_impact_2004} Visceral obesity rather than body mass index has been reported to be associated with greater incidence of wound complications, overall complication rate and prolonged hospital stay in pateints undergoing laparascopic sigmoid colectomy for cancer. \parencite{tsujinaka_visceral_2008}

However, in a prospective, multi-institutional, risk-adjusted cohort study of 2258 patients including 699 patients who underwent pancreatectomy for cancer, Mullen and co-workers reported that obesity on its own was not a predictor of major postoperative complications or mortality. However, they noted that underweight patients had a five-fold increased risk of postoperative mortality. They also reported that obesity was an independent predictor of minor complications including wound infections.\parencite{mullen_impact_2008}

Dindo and co-workers studied the impact of obesity on postoperative outcomes in a prospective study of 6336 patients undergoing elective general surgery. They reported that obesity was not an independent risk factor for morbidity or mortality. There was no difference in the median operating time, blood transfusions or lenght of hospital stay between obese and non-obese patients. The type and severity of complications were similar between obese and non-obese patients. \parencite{dindo_obesity_2003}

In a more recent prospective, multi-institutional, risk-adjusted cohort study of 118,707 patients undergoing non-bariatric general surgery, Mullen co-workers reported that the risk of mortality was lowest in the overweight (BMI 25 - 30) and moderately obese (BMI 30 - 35) patients and highest in the underweight and morbidly obese patients. This appears to confirm the existence of an 'obesity paradox' in surgical patients. \parencite{mullen_obesity_2009}
